Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-03-01 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
hs: more refs

%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{figure}
\begin{tikzpicture}[baseline,>=stealth]
    % \draw[help lines, gray] (-4, -2) grid (4, 2);
    \draw[->] (-2.5, 0) -- (-1.75, +0.5);
    \draw[->] (-2.5, 0) -- (-1.75, -0.5);

    \node at (-3.25, 0) {$\{h_{\alpha\beta}, \, \varphi\}$};

    \node at (-1, 1.25) {${\cal O}(\chi^{0})$};
    \node at (-1, +0.5) {$\{ \textcolor{rred}{h^{\rm p}_{\alpha\beta}}, \, \varphi\}$};
    \node at (-1, -0.5) {$h^{\rm a}_{\alpha\beta}$};

    \node at (+1, 1.25) {${\cal O}(\chi^{1})$};
    \node at (+1, +0.5) {\dots};
    \node at (+1, -0.5) {\dots};
\end{tikzpicture}
\caption{Quasinormal modes of BHs in scalar-Gauss-Bonnet gravity.}
\end{figure}

@@

\begin{figure}
\begin{tikzpicture}[baseline,>=stealth]
    % \draw[help lines, gray] (-4, -2) grid (4, 2);
    \draw[->] (-2.5, 0) -- (-1.75, +0.5);
    \draw[->] (-2.5, 0) -- (-1.75, -0.5);

    \node at (-3.25, 0) {$\{h_{\alpha\beta}, \, \vartheta\}$};

    \node at (-1, 1.25) {${\cal O}(\chi^{0})$};
    \node at (-1, +0.5) {$h^{\rm p}_{\alpha\beta}$};
    \node at (-1, -0.5) {$\{ \textcolor{rred}{h^{\rm a}_{\alpha\beta}}, \, \vartheta\}$};

    \node at (+1, 1.25) {${\cal O}(\chi^{1})$};
    \node at (+1, +0.5) {\dots};
    \node at (+1, -0.5) {\dots};
\end{tikzpicture}
\caption{Quasinormal modes of BHs in dynamical Chern-Simons gravity.}
\end{figure}

@@

We can now expand Eq.~\eqref{eq:omega_dsc_eg} to leading-orders in $\chi$ and $\ell_{\rm CS}$, and
gather the terms proportional to $\ell_{\rm CS}$.
%
Doing so, we find,
%
\begin{align}
    \Mf \, \omega^{\rm CS}_{220} &=
    \left[ 0.3722 + 1.1945 (\ell_{\rm CS} / \Mf)^4 \right]
    \nonumber \\
    &\quad + \left[0.1861 + 5.1828 (\ell_{\rm CS} / \Mf)^4 \right] \, \chi\,,
\end{align}
%
where we made use of the numerical value of the coefficients $c_{i}$ and,
reassuringly, the nonrotating and linear-in-spin coefficients above agree with those
of~\cite{Maselli:2019mjd} to xx\%.
%
We now isolate the dCS corrections and compare against Eq.~\eqref{eq:qnm_nongr_iso}.
%
Doing so, we can make the identifications,
%
\begin{equation}
p_{\rm CS} = 4\,, \quad \delta \omega^{(0)}_{\rm CS} = 3.1964\,, \quad \delta \omega^{(1)}_{\rm CS} = 41.199\,.
\label{eq:cs_omega_coefs}
\end{equation}
%
The same procedure can be carried for $\tau^{(i)}$, which results in,
%
\begin{equation}
\delta \tau^{(0)}_{\rm CS} = 6.3619\,, \quad \delta \tau^{(1)}_{\rm CS} = 794.66\,.
\label{eq:cs_tau_coefs}
\end{equation}
%
which completes the set of fixed parameters beyond-GR parameters in the ringdown of the \pSEOB{} waveform model.
%
As we said before, the only free beyond-GR parameter left is now $\ell_{\rm CS}$.

@@

\begin{table*}[th]
\begin{tabular}{c | c c c c c c c c}
\hline
\hline
Theory & $p$ & $\delta \omega^{(0)}_{220}$ & $\delta \tau^{(0)}_{220}$ & $\delta \omega^{(1)}_{220}$ & $\delta \tau^{(1)}_{220}$ & QNM & Constraint & This work \\
\hline
sGB      & 4 & 0.0107 & 0.0044 & \dots & \dots &  \cite{Pierini:2021jxd} & $\ell_{\rm GB} \leqslant 1.18$~km~(GW)~\cite{Lyu:2022gdr} & \dots \\
EdGB      & 4 & 3.1964 & 6.3619 & \dots & \dots &  \cite{Wagle:2021tam}   & $\ell_{\rm CS} \leqslant 8.5$~km~(EM+GW)~\cite{Silva:2020acr}  & $\ell_{\rm CS} \leqslant 34.5$~km \\
cEFTofGR & 4 & $-0.3680$ $(-0.5131)$  & 1.9242 (1.7000) & \dots & \dots & \cite{Cano:2021myl} & --  & \dots \\
qEFTofGR & 6 & \dots & \dots & \dots & \dots & \cite{Cano:2021myl} & --  & \dots \\
\hline
\hline
\end{tabular}
\caption{Summary of the quasinormal modes calculations.
%
We summarize each theory we have considered together with: the exponent $p$ at
which their QNM-modification enters, the corresponding modifications to the
oscillation frequency $\delta \omega^{(i)}_{220}$ and decay time $\delta \tau^{(i)}_{220}$, the
references from which we used the results from and the current best constraint
(if applicable).
%
In the entries for cEFTofGR, the first entries in the $\delta
\omega^{(i)}_{220}$, $\delta \tau^{(i)}_{220}$ are obtained by considering
only parity-even corrections and the second entries by considering
only parity-odd corrections.
%
In these cases, the dimensionless parameters $\lambda_{\rm e,o}$ are degenerate
with the length scale $\ell$.
%
}
\label{tab:ref_theories_qnms}
\end{table*}

@@

Among our beyond-GR parameters, $\{\ell, \delta \omega_k^{(j)}\delta \tau_k^{(j)}\}$, as already mentioned in the previous section, we would mostly hold $\{\delta \omega_k^{(j)}\},$ and $\{\delta \tau_k^{(j)}\}$ fixed to theory-specific predictions allowing only $\ell$ to freely vary. For such cases, we assume a prior uniform between appropriate ranges on $\ell$.

@@

When possible, we also combine the marginalized posterior distribution on
$\ell$ for both event GW150914 and GW200129. To do so, we employ the same
approach as Ref.~\cite{Perkins:2021mhb} which relies in obtaining a kerneld
density estimator (KDE) as an approximation for the posterior distribution on
$\ell$ for each individual events which are then multiplied together to yield
the cumulative posterior.
%
This method can have difficulties in dealing with hard cut-offs in the
distribution, such as the one we impose at $\ell = 0$. We address this issue as
done in Ref.~\cite{Perkins:2021mhb}, by first doubling our number of samples as
$\{ s_i \} \cup \{ - s_i \}$, i.e., mirroring the samples across $\ell = 0$.
%
We then renormalize the associated KDE numerically in the range $\ell = [0,
\infty]$.

@@

@@

We also found for both GW events, that the
perturbative-conditions~\eqref{eq:parspec_bound} required by the {\sc Parspec}
formalism is violated for the $\gamma_{\rm CS} \delta \tau^{(1)}$ coefficient.
%
This means that we cannot use these posterior to infer any meaningful bound on dCS gravity
and that is why we quoted only the $j=0$ bound above.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% General
% Motivation
% What we do
% Main conclusion
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%

@@

% \begin{figure}
% \begin{tikzpicture}[baseline,>=stealth]
%     % \draw[help lines, gray] (-4, -2) grid (4, 2);
%     \draw[->] (-2.5, 0) -- (-1.75, +0.5);
%     \draw[->] (-2.5, 0) -- (-1.75, -0.5);
%
%     \node at (-3.25, 0) {$\{h_{\alpha\beta}, \, \varphi\}$};
%
%     \node at (-1, 1.25) {${\cal O}(\chi^{0})$};
%     \node at (-1, +0.5) {$\{ \textcolor{rred}{h^{\rm p}_{\alpha\beta}}, \, \varphi\}$};
%     \node at (-1, -0.5) {$h^{\rm a}_{\alpha\beta}$};
%
%     \node at (+1, 1.25) {${\cal O}(\chi^{1})$};
%     \node at (+1, +0.5) {\dots};
%     \node at (+1, -0.5) {\dots};
% \end{tikzpicture}
% \caption{Quasinormal modes of BHs in scalar-Gauss-Bonnet gravity.}
% \end{figure}

@@

% \begin{figure}
% \begin{tikzpicture}[baseline,>=stealth]
%     % \draw[help lines, gray] (-4, -2) grid (4, 2);
%     \draw[->] (-2.5, 0) -- (-1.75, +0.5);
%     \draw[->] (-2.5, 0) -- (-1.75, -0.5);
%
%     \node at (-3.25, 0) {$\{h_{\alpha\beta}, \, \vartheta\}$};
%
%     \node at (-1, 1.25) {${\cal O}(\chi^{0})$};
%     \node at (-1, +0.5) {$h^{\rm p}_{\alpha\beta}$};
%     \node at (-1, -0.5) {$\{ \textcolor{rred}{h^{\rm a}_{\alpha\beta}}, \, \vartheta\}$};
%
%     \node at (+1, 1.25) {${\cal O}(\chi^{1})$};
%     \node at (+1, +0.5) {\dots};
%     \node at (+1, -0.5) {\dots};
% \end{tikzpicture}
% \caption{Quasinormal modes of BHs in dynamical Chern-Simons gravity.}
% \end{figure}

@@

We can now expand Eq.~\eqref{eq:omega_dsc_eg} to leading-orders in $\chi$ and $\ell_{\rm CS}$, and
gather the terms proportional to $\ell_{\rm CS}$.
%
Doing so, we find,
%
\begin{align}
    \Mf \, \omega^{\rm CS}_{220} &=
    \left[ 0.3722 + 1.1945 (\ell_{\rm CS} / \Mf)^4 \right]
    \nonumber \\
    &\quad + \left[0.1861 + 5.1828 (\ell_{\rm CS} / \Mf)^4 \right] \, \chi\,,
\end{align}
%
where we made use of the numerical value of the coefficients $c_{i}$ and,
reassuringly, the nonrotating and linear-in-spin coefficients above agree with those
of~\cite{Maselli:2019mjd} to xx\%.
%
We now isolate the dCS corrections and compare against Eq.~\eqref{eq:qnm_nongr_iso}.
%
Doing so, we can make the identifications,
%
\begin{equation}
p_{\rm CS} = 4\,, \quad \delta \omega^{(0)}_{\rm CS} = 3.1964\,, \quad \delta \omega^{(1)}_{\rm CS} = 41.199\,.
\label{eq:cs_omega_coefs}
\end{equation}
%
The same procedure can be carried for $\tau^{(i)}$, which results in,
%
\begin{equation}
\delta \tau^{(0)}_{\rm CS} = 6.3619\,, \quad \delta \tau^{(1)}_{\rm CS} = 794.66\,.
\label{eq:cs_tau_coefs}
\end{equation}
%
which completes the set of fixed parameters beyond-GR parameters in the ringdown of the \pSEOB{} waveform model.
%
We remark that the alarmingly large values of $\delta \omega_{\rm CS}^{(1)}$ and $\delta \tau_{\rm CS}^{(1)}$
are compensated by the assumptions that $(\ell_{\rm CS} / M_f)^2$ and $\chi$ are much less than unity. These
are the assumptions use to calculate the QNM frequencies both in Ref.~\cite{Wagle:2021tam} and also in
other gravity considered here~\cite{Pierini:2021jxd,Cano:2021myl}.
%
As we said before, the only free beyond-GR parameter left is now $\ell_{\rm CS}$.

@@

\begin{table*}[th]
\begin{tabular}{c | c c c c c c c c}
\hline
\hline
Theory & $p$ & $\delta \omega^{(0)}_{220}$ & $\delta \tau^{(0)}_{220}$ & $\delta \omega^{(1)}_{220}$ & $\delta \tau^{(1)}_{220}$ & QNM & Constraint & This work \\
\hline
EdGB      & 4 & 0.0107 & 0.0044 & \dots & \dots &  \cite{Pierini:2021jxd} & $\ell_{\rm GB} \leqslant 1.18$~km~(GW)~\cite{Lyu:2022gdr} & \dots \\
dCS      & 4 & 3.1964 & 6.3619 & \dots & \dots &  \cite{Wagle:2021tam}   & $\ell_{\rm CS} \leqslant 8.5$~km~(EM+GW)~\cite{Silva:2020acr}  & $\ell_{\rm CS} \leqslant 34.5$~km \\
cEFTofGR & 4 & $-0.3680$ $(-0.5131)$  & 1.9242 (1.7000) & \dots & \dots & \cite{Cano:2021myl} & --  & \dots \\
% qEFTofGR & 6 & \dots & \dots & \dots & \dots & \cite{Cano:2021myl} & --  & \dots \\
\hline
\hline
\end{tabular}
\caption{Summary of the quasinormal modes calculations.
%
We summarize each theory we have considered together with: the exponent $p$ at
which their QNM-modification enters, the corresponding modifications to the
oscillation frequency $\delta \omega^{(i)}_{220}$ and decay time $\delta \tau^{(i)}_{220}$, the
references from which we used the results from and the current best constraint
(if applicable).
%
In the entries for cEFTofGR, the first entries in the $\delta
\omega^{(i)}_{220}$, $\delta \tau^{(i)}_{220}$ are obtained by considering
only parity-even corrections and the second entries by considering
only parity-odd corrections.
%
In these cases, the dimensionless parameters $\lambda_{\rm e,o}$ are degenerate
with the length scale $\ell$.
%
}
\label{tab:ref_theories_qnms}
\end{table*}

@@

Among our beyond-GR parameters, $\{\ell, \delta \omega_k^{(j)}\delta \tau_k^{(j)}\}$, as already mentioned in the previous section, we would mostly hold $\{\delta \omega_k^{(j)}\},$ and $\{\delta \tau_k^{(j)}\}$ fixed to theory-specific predictions allowing only $\ell$ to freely vary. For such cases, we assume a prior uniform between appropriate ranges on $\ell$, making sure that our marginalized posterior distributions on $\ell$ do not rail over the maximum prior value on $\ell$.

@@

@@

If we find that our posterior obey both conditions~\eqref{eq:eft_bound}
and~\eqref{eq:parspec_bound}, we combine individual posteriors on $\ell$ from
each event.
%
% When possible, we also combine the marginalized posterior distribution on
% $\ell$ for both event GW150914 and GW200129.
%
To do so, we employ one of the approaches of Ref.~\cite{Perkins:2021mhb} which
is based on obtaining a kernel density estimator (KDE) as an approximation for
the posterior distribution on $\ell$ for each individual events. These are then
multiplied together to yield the cumulative posterior.
%
(See also Refs.~\cite{DelPozzo:2011pg,Cardenas-Avendano:2019zxd,Carullo:2021dui}.)
%
This method can have difficulties in dealing with hard cut-offs in the
distribution, such as the one we impose at $\ell = 0$. We address this issue as
done in Ref.~\cite{Perkins:2021mhb}, by first doubling our number of samples as
$\{ s_i \} \cup \{ - s_i \}$, i.e., mirroring the samples across $\ell = 0$.
%
We then renormalize the associated KDE numerically in the range $\ell = [0,
\infty]$.

@@

We also found for both GW events, that the
perturbative-conditions~\eqref{eq:parspec_bound} required by the {\sc Parspec}
formalism is violated for the $\gamma_{\rm CS} \, \delta \tau^{(1)}$ coefficient.
%
This means that we cannot use these posterior to infer any meaningful bound on dCS gravity
and that is why we quoted only the $j=0$ bound above.
